### 2023-01-14 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@b47a6b7b3faa32128c1c7850dd6a20efef0eca20 🚀

Why is this useful? For one, this analytic method avoids the extra
computation required by simulating samples from this distribution
directly. Further, it enables continuous exanination of the number of
particles \(n_p\) as a uncertainty width. The natural analogue for the
number of particles sampled is the number of observations of
co-occurrence for a given pair of species---the fewer the particles, the
higher the variance of the resulting approximation (see supplemental
figure 1 for an example). Equation (\textbf{ref:eq:eq1?}) is undefined
for 0 particles (i.e.~0 observations co-occurrence), although as \(n_p\)
approaches \(0\) the approxated normal (once truncated) approaches a
uniform distribution on the interval \((0,1)\), the maximum entropy
distribution where we have no information about the possibility of an
interaction.

This also has implications for what we mean by `uncertainty' in
interaction predictions. A model's prediction can be `uncertain' in two
different ways: (1) the model's predictions may have high variance, or
(2) the model's predictions may be centered around a probability of
interaction of \(0.5\), where we are the most unsure about whether this
interaction exists . For non-zero \(p_{fn}\) and \(p_{fp}\), the above
model always moves the mean interaction probability toward 0.5
(supplemental figure 2), siding toward higher uncertainty in the latter
(higher entropy) sense. Improving the incorporation of different forms
of uncertainty in probabilistic interaction predictions seems a
necessary next step toward understanding what pairs of species we know
the least about, in order to prioritize sampling to provide the most new
information possible.

@@

The model above doesn't consider the possibility that there are positive
or negative associations which shift the probability of species
cooccurrence together due to their interaction (Cazelles \emph{et al.}
2016). However, here we demonstrate that the probability of having a
false-negative can be higher if there is some positive association in
the occurrence of species \(A\) and \(B\). If we denote the probability
that we observe the co-occurrence of two species \(A\) and \(B\) as
\(P(AB)\) and if there is no association between the marginal
probabilities of observing \(A\) and observing \(B\), denoted \(P(A)\)
and \(P(B)\) respectively, then the probability of observing their
co-occurrence is the product of the marginal probabilities for each
species, \(P(AB) = P(A)P(B)\). In the other case where there is some
positive strength of association between observing both \(A\) and \(B\)
because this interaction is ``important'' for each species, then the
probability of observation both \(A\) and \(B\), \(P(AB)\), is greater
than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent and
instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false-negative-rate, as the
probability of false negative would not be constant for each pair of
species across sites.

>>>>

Why is this useful? For one, this analytic method avoids the extra
computation required by simulating samples from this distribution
directly. Further, it enables continuous exanination of the number of
particles \(n_p\) as a uncertainty width. The natural analogue for the
number of particles sampled is the number of observations of
co-occurrence for a given pair of species---the fewer the particles, the
higher the variance of the resulting approximation. The normal
approximation is undefined for 0 particles (i.e.~0 observations
co-occurrence), although as \(n_p\) approaches \(0\) the approxated
normal (once truncated) approaches a uniform distribution on the
interval \((0,1)\), the maximum entropy distribution where we have no
information about the possibility of an interaction.

This also has implications for what we mean by `uncertainty' in
interaction predictions. A model's prediction can be `uncertain' in two
different ways: (1) the model's predictions may have high variance, or
(2) the model's predictions may be centered around a probability of
interaction of \(0.5\), where we are the most unsure about whether this
interaction exists. Improving the incorporation of different forms of
uncertainty in probabilistic interaction predictions seems a necessary
next step toward understanding what pairs of species we know the least
about, in order to prioritize sampling to provide the most new
information possible.

@@

The model above doesn't consider the possibility that there are positive
or negative associations which shift the probability of species
cooccurrence away from what is expected based on their relative
abundances due to their interaction (Cazelles \emph{et al.} 2016).
However, here we demonstrate that the probability of having a
false-negative can be higher if there is some positive association in
the occurrence of species \(A\) and \(B\). If we denote the probability
that we observe the co-occurrence of two species \(A\) and \(B\) as
\(P(AB)\) and if there is no association between the marginal
probabilities of observing \(A\) and observing \(B\), denoted \(P(A)\)
and \(P(B)\) respectively, then the probability of observing their
co-occurrence is the product of the marginal probabilities for each
species, \(P(AB) = P(A)P(B)\). In the other case where there is some
positive strength of association between observing both \(A\) and \(B\)
because this interaction is ``important'' for each species, then the
probability of observation both \(A\) and \(B\), \(P(AB)\), is greater
than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent and
instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false-negative-rate, as the
probability of false negative would not be constant for each pair of
species across sites.
